The Soulgem Paper

### 2011-09-22 soulgem.tex
First draft of Discussion - Summary.

@@

% First a brief summary of the findings

% Second, state that the main theoretical foundations held, and were improved upon

>>>>

% TODO: Need to add *bolded* sentences to point out all of the findings in each section.

@@

We explored the sociological and organizational literature to generate a theoretical framework of roles and expectations in software organizations. Using that framework we studied several research questions regarding work in software groups.

We found that roles, defined as sets of expectations, were inconsistent within and between the organizations we studied. Roles are socially constructed; part of the construction occurs at the cultural or institutional level, and part occurs at the local, team level, and is caused by historical, structural, and personal factors. Therefore, although we can make some reasonable assumptions about the work of a software professional based on his role label or official position, these assumptions will only take us so far: professionals vary considerably in their responsibilities, and treating them as archetypical professionals (``developers,'' ``testers,'' and so on) dismisses a large part of what they bring or fail to bring to their job.

We formulated a list of expectations that were common or significant in the two sites we observed. We do not think that our list is exhaustive, but we do think that it represents expectations of concern in most software organizations, and that formulating such lists and analyzing how expectations map to organization members gives us a good understanding of role definitions and interactions for software development.

We also explored role evolution and found that it is driven by organizational growth (which leads to specialization) and by mismatches between expectations and behaviour (which lead to a readjustment of expectations). Finally, we found that people enacting different roles tend to use different media to communicate, and that face-to-face interaction and issue-tracking notifications were the most widely used media in our sites in terms of inter-role communications, with the qualification that issue-tracking systems were used differently by people enacting different roles.
